---
title: "Aspect importance - demo"
author: "Katarzyna Pękala"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Aspect importance - demo}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```


# Demo

## Aspect importance

Aspect importance function provides instance-level explanations for the groups of explanatory variables. It enables grouping predictors into entities called aspects. Afterwards, it can calculate the contribution of those aspects to the prediction.

To illustrate how the function works, we use titanic example. We build random 
forest model, group features into aspects and choose new observation to be explained. Then we build `DALEX` explainer and use it to call aspect importance function. Finally, we print and plot function results. We can observe that `personal` (`age` and `gender`) variables have the biggest contribution to the prediction. This contribution is of a positive type.

```{r}
library("DALEX")
library("randomForest")
library("DALEXtra")
titanic <- titanic_imputed
titanic$country <- NULL
titanic_without_target <- titanic[,colnames(titanic)!="survived"]

aspects_titanic <-
  list(
    wealth = c("class", "fare"),
    family = c("sibsp", "parch"),
    personal = c("age", "gender"),
    embarked = "embarked"
  )

passenger <- titanic_without_target[4,]

model_titanic_rf <- randomForest(factor(survived) == "yes" ~ gender + age + 
                                   class + embarked + fare + sibsp + parch,  
                                 data = titanic)

predict(model_titanic_rf, passenger)
```


```{r}
explain_titanic_rf <- explain(model_titanic_rf, 
                               data = titanic_without_target,
                               y = titanic$survived == "yes", 
                               predict_function = predict,
                               verbose = FALSE)

titanic_rf_ai <- aspect_importance(x = explain_titanic_rf, 
                                   new_observation = passenger, 
                                   aspects = aspects_titanic)
titanic_rf_ai

plot(titanic_rf_ai, add_importance = TRUE)

```


## Automated grouping

In examples described above, we had to manually group features into aspects. 
Aspect importance provides group_variables() - function that automatically groups features for us, based on the features correlation. Function only works on numeric variables.  

Below, we test aspect importance function on another dataset. But this time we build aspect list by running run group_variables() (with cut off level set on 0.6). As a result, we get a list of variables groups (aspects) where absolute value of features' pairwise correlation is at least at 0.6.


```{r import apartments}
library(DALEX)
data("apartments")
apartments_num <- apartments[,unlist(lapply(apartments, is.numeric))] 
apartments_no_target <- apartments_num[,-1]
new_observation_apartments <- apartments_num[10,]
model_apartments <- lm(m2.price ~ ., data = apartments_num)
aspects_apartments <- group_variables(apartments_no_target, 0.6)
predict(model_apartments, new_observation_apartments)

```


```{r}
explain_apartments <- explain(model_apartments, 
                               data = apartments_no_target,
                               y = apartments$m2.price, 
                               predict_function = predict,
                               verbose = FALSE)
apartments_ai <- aspect_importance(x = explain_apartments, 
                                   new_observation = new_observation_apartments, 
                                   aspects = aspects_apartments, 
                                   N = 1000, show_cor = TRUE)
apartments_ai
plot(apartments_ai, aspects_on_axis = FALSE, add_importance = TRUE, 
     digits_to_round = 0)
```

## Triplot

`Triplot` is one more tool that allows us to better understand the inner workings a of black box model. After calling triplot and providing it with model, dataset and  observation to be explained, it illustrates (in one place): 

* single aspect importance - aspect importance results when every aspect has only one feature,
* hierarchical aspects importance (description can be found in vignette), 
* order of grouping features into aspects in `group_variables()`.

```{r}
triplot(model_apartments, apartments_no_target, 
        new_observation = new_observation_apartments, N = 5000, 
        add_importance_labels = FALSE, axis_lab_size = 8, abbrev_labels = 11)

```
